---
title: "Project 3: PROJECT3PACKAGE Tutorial"
author: "Sean Grimm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PROJECT3PACKAGE Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PROJECT3PACKAGE)
```

# my_t.test

The following shows example uses of ```my_t.test``` . Using the ```my_gapminder``` data, we calculate the mean life expectancy in this data set is about 59.47 years. We will conduct three t-tests to assess the validity of the null hypothesis $\{H_{0}}$ = 60.

```{r}
life_expectancy <- my_gapminder[["lifeExp"]]
mean(life_expectancy)
```

```{r}
my_t.test(life_expectancy, "two.sided", 60)
```

In this test, we see that the returned p-value is about 0.093. Since we used ```alternative = "two.sided"```, this precisely means that, assuming the mean life expectancy is indeed 60, there is a probability of 0.093 that another sample of the exact same size taken under the exact same conditions will be as or more extreme as the mean life expectancy of our data set. Using the significance level $\alpha$ = 0.05, we would fail to reject the null hypothesis. As such, we cannot conclude that the mean life expectancy from our data set is not 60.

```{r}
my_t.test(life_expectancy, "less", 60)
```

This t-test yields a p-value of about 0.046. Since the p-value is less than our significance level $\alpha$ = 0.05, we would reject the null hypothesis in this case, thus concluding that the mean life expectancy of the data set is, in actuality, less than 60.

```{r}
my_t.test(life_expectancy, "greater", 60)
```

Lastly, this t-test yields a p-value of about 0.953. This is far larger than our significance level $\alpha$ = 0.05, and as such, we would fail to reject the null hypothesis. Once again, we would not be able to conclude that the mean life expectancy for the data set is not 60.
